---
title: "file-validation"
author: "Alexander Nielson"
date: "3/4/2020"
output: html_document
---

#Description
allows you to check an uploaded file that has not yet been processed.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
options(scipen=999)
library(tidyverse)
library(readxl)
library(lubridate)
library(odbc)
```

```{r}
file_path <- "E:\\rprojects\\entity-debugging-toolkit\\downloaded-batches\\wallace-stegner\\_data_64596.txt"

file_data<- read_delim(file_path, delim = "|", col_names = FALSE)

names(file_data) <- c("record_number",                                 
                      "fiscal_year",                                   
                      "entity_name",                                   
                      "fund1",                                           
                      "fund2",
                      "fund3",
                      "fund4",
                      "o1",
                      "o2",
                      "o3",
                      "o4",
                      "o5",
                      "o6",
                      "o7",
                      "o8",
                      "o9",
                      "o10",
                      "transaction_type",
                      "cat1",
                      "cat2",
                      "cat3",
                      "cat4",
                      "cat5",
                      "cat6",
                      "cat7",
                      "pcv_name",
                      "payee_dba_name",
                      "vendor_id_code",
                      "posting_date",
                      "transaction_description",
                      "transaction_id",
                      "transaction_reference_id",
                      "contract_name",
                      "contract_number",
                      "position_title",
                      "hourly_rate",
                      "gender",
                      "amount",
                      "protection_indicator"#,
                      #"uca_block",
                      # "p1",
                      # "p2",
                      # "p3",
                      # "p4",
                      # "p5",
                      # "p6",
                      # "p7",
                      # "funct1",
                      # "funct2",
                      # "funct3",
                      # "funct4",
                      # "funct5",
                      # "funct6",
                      # "funct7"
                      )
                

```
#Common check:

```{r}
file_data %>% summarize(
  net_amount = paste("$",formatC(sum(amount), format="d", big.mark=",")),
  num_of_records = formatC(n(),format="d", big.mark=",")
)
```



#validate record number
```{r}
if(file_data %>% 
  group_by(record_number) %>%
  summarize(n=n()) %>% 
  filter(n>1) %>%
  nrow() > 0){
    stop("duplicate 'record number' observations")
  }
```

#validate fiscal year
```{r}
# if(file_data %>% 
#   group_by(record_number) %>%
#   summarize(n=n()) %>% 
#   filter(n>1) %>% nrow() > 0){
#     stop("duplicate 'record number' observations")
#   }
```

# entity name
```{r}
if(file_data %>% 
  filter(nchar(entity_name) > 60) %>%
  nrow() > 0){
    stop("records with entity name too long > 60 characters")
  }
```

#fund1 1 nam
```{r}
if(file_data %>% 
  filter(nchar(fund1) > 60) %>%
  nrow() > 0){
    stop("records with fund 1 name too long > 60 characters")
}

if(file_data %>% 
  filter(is.na(fund1)) %>%
  nrow() > 0){
    stop("records with missing fund1 cells")
  }
```

# organization level 1 name
```{r}
if(file_data %>% 
  filter(nchar(o1) > 60) %>%
  nrow() > 0){
    stop("records with organization level 1 name too long > 60 characters")
}

if(file_data %>% 
  filter(is.na(o1)) %>%
  nrow() > 0){
    stop("records with missing organizaiton level 1 cells")
  }
```


# transaction type
```{r}
if(file_data %>% 
  filter(nchar(transaction_type) > 2) %>%
  nrow() > 0){
    stop("records with a transaction type > 2")
}

if(file_data %>%
   filter(
     transaction_type != "EX" &&
     transaction_type != "PY" &&
     transaction_type != "RV" &&
     transaction_type != "BG" &&
     transaction_type != "BS"
   ) %>% nrow() > 0) {
  stop("records with an invalid transaction type ")
}

```

# category level 1 name
```{r}
if(file_data %>% 
  filter(nchar(cat1) > 60) %>%
  nrow() > 0){
    stop("records with category level 1 name too long > 60 characters")
}

if(file_data %>% 
  filter(is.na(cat1)) %>%
  nrow() > 0){
    stop("records with missing category level 1 cells")
  }
```


# pcv_name
```{r}
if(file_data %>% 
  filter(nchar(pcv_name) > 60) %>%
  nrow() > 0){
    stop("records with payee/customer/vendor name too long > 60 characters")
}

if(file_data %>% 
  filter(is.na(pcv_name)) %>%
  nrow() > 0){
    stop("records with missing payee/customer/vendor name cells")
  }
```


# vendor_id_code
```{r}
if(file_data %>% 
  filter(nchar(vendor_id_code) > 60) %>%
  nrow() > 0){
    stop("records with vendor_id_code name too long > 60 characters")
}

if(file_data %>% 
  filter(is.na(vendor_id_code)) %>%
  nrow() > 0){
    stop("records with missing vendor_id_code cells")
  }
```

# posting_date
```{r}
if(file_data %>% 
  filter( !is.POSIXct(posting_date) ) %>%
  nrow() > 0){
  
  file_data <- file_data%>% mutate(posting_date = mdy(posting_date))
    #stop("records with incorrect posting_date format")
}


```

# transaction_description
```{r}
if(file_data %>% 
  filter(nchar(transaction_description) > 255) %>%
  nrow() > 0){
    stop("records with transaction_description name too long > 255 characters")
}

if(file_data %>% 
  filter(is.na(transaction_description)) %>%
  nrow() > 0){
    stop("records with missing transaction_description cells")
  }


```


# transaction_id
```{r}
if(file_data %>% 
  filter(nchar(transaction_id) > 255) %>%
  nrow() > 0){
    stop("records with transaction_id name too long > 255 characters")
}

if(file_data %>% 
  filter(is.na(transaction_id)) %>%
  nrow() > 0){
    stop("records with missing transaction_id cells")
  }


```

# contract_name
```{r}
if(file_data %>% 
  filter(nchar(contract_name) > 255) %>%
  nrow() > 0){
    stop("records with contract_name name too long > 255 characters")
}

```

# contract_number
```{r}
if(file_data %>% 
  filter(nchar(contract_number) > 50) %>%
  nrow() > 0){
    stop("records with contract_number name too long > 50 characters")
}
```

# position_title
```{r}
if(file_data %>% 
  filter(nchar(position_title) > 50) %>%
  nrow() > 0){
    stop("records with contract_name name too long > 50 characters")
}

```


# hourly_rate
```{r}
if(file_data %>% 
  filter(!is.numeric(hourly_rate) ) %>%
  nrow() > 0){
    stop("records with hourly_rate incorrect format")
}

```


# gender
```{r}
if(file_data %>% 
  filter(nchar(gender) > 1 ) %>%
  nrow() > 0){
    stop("records with gender too long > 1 characters")
}

if(file_data %>% 
  filter(gender != "M" &&
         gender != "F") %>%
  nrow() > 0){
    stop("records with gender incorrect format. must be M or F")
}

```

# amount
```{r}
net_amount <- file_data %>% pull(amount) %>% sum()
print(net_amount)
```

